Development of a Python Package for Matching Observational Data

### 2022-05-06 paper/main.tex
more small edits

Another valid distance measure is the $L^p$ norms of the difference between feature vectors, i.e. $\| \vec{x}_T - \vec{x}_C \|_p$. 
As~\textcite{king_why_2019} note, norm-based distance metrics attempt to replicate fully-blocked design, and thus offer some theoretical advantages.
However, they do not incorporate any weighting about a feature's ``importance'', as we saw with balancing scores and almost exact matching.
Hence, norms are most often used in the form of Mahalanobis distance, which first rescales all features to have mean zero and unit variance, and then calculates distances using the $L^2$ norm.
This ensures that all features have equal weight in the matching, which may or may not be desirable; in either case, it almost certainly better than the no-rescaling case, where a feature's variance would determine its relative importance.

@@

\textcite{baltar_mahalanobis_2014} explored the efficacy of this iterative method in practice, relative to propensity score and Mahalanobis distance alone. In the paper, they compare seek to estimate the effect of Brazilian towns' social agenda on public health. They found both the propensity score alone and iterative method to perform admirably, whereas Mahalanobis distance performed poorly; they do not take this as an indictment of Mahalanobis distance matching, and pose for further research to identify conditions under which Mahalnobis distance may perform \emph{better}. The iterative method achieved the greatest reduction in bias for all features, except the propensity score, for which propensity score distance alone (somewhat obviously) saw the greatest bias reduction. However, as~\textcite{austin_optimal_2011} and~\textcite{greifer_assessing_2022} note, balance of the scores does not imply balance of the features, and balancing score balance metrics should only be used as a supplementary diagnostic; the focus of balance assessment should be balance of the features.

@@

\textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply the Munkres algorithm~\parencite{munkres_algorithms_1957}. We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observations, and $j$ a control-group observations. The Hungarian algorithm solves the \emph{linear assignment problem}, which in this setting may be expressed as
\begin{equation}
  \min_{\set{M}} \sum_{(i, j) \in \set{M}} d(\vec{x}_{T_{i}}, \vec{x}_{C_{j}}), \text{ such that } \forall (i, j), (k, l) \in \set{M}, i = k \iff j = l;
\end{equation}
the ``such that'' condition simply specifies that the matching is one-to-one. As such, optimal matching via the Hungarian algorithm minimizes the total sum of distances between matches in the matched subset.

\paragraph{Extensions.} The Hungarian algorithm can be extended to the $1:k$ matching case via $b$-matching while retaining optimality.
This process involves adding $k$ ``copy'' nodes for each node representing a $T$ datapoint, each with identical edges as the original node, and then perform the optimal matching on the resulting graph.
However, as~\textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases.~\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.

>>>>

A rather obvious distance measure is the $L^p$ norm of the difference between feature vectors, i.e. $\| \vec{x}_T - \vec{x}_C \|_p$. 
As~\textcite{king_why_2019} note, norm-based distance metrics attempt to replicate fully-blocked design, and thus offer some theoretical advantages.
However, they do not incorporate any weighting about a feature's ``importance'', as we saw with balancing scores and almost exact matching.
Hence, norms are most often used in the form of Mahalanobis distance, which first rescales all features to have mean zero and unit variance, and then calculates distances using the $L^2$ norm.
This ensures that all features have equal weight in the matching, which may or may not be desirable; in either case, it almost certainly better than the no-rescaling case, where a feature's variance would determine its relative importance.

@@

\textcite{baltar_mahalanobis_2014} explored the efficacy of this iterative method in practice, relative to propensity score and Mahalanobis distance alone. In the paper, they seek to estimate the effect of Brazilian towns' social agenda on public health. They found both the propensity score alone and iterative method to perform admirably, whereas Mahalanobis distance performed poorly; they do not take this as an indictment of Mahalanobis distance matching, and pose for further research to identify conditions under which Mahalnobis distance may perform \emph{better}. The iterative method achieved the greatest reduction in bias for all features, except the propensity score, for which propensity score distance alone (somewhat obviously) saw the greatest bias reduction. However, as~\textcite{austin_optimal_2011} and~\textcite{greifer_assessing_2022} note, balance of the scores does not imply balance of the features, and balancing score balance metrics should only be used as a supplementary diagnostic; the focus of balance assessment should be balance of the features.

@@

\textcite{rosenbaum_optimal_1989} argues for optimal matching according to the Hungarian algorithm, also known as the Kuhn-Munkres or simply the Munkres algorithm~\parencite{munkres_algorithms_1957}. We will denote a matching a set $\set{M}$ of tuples $(i, j)$, with $i$ representing a treatment-group observation $T_i$, and $j$ a control-group observation $C_j$. The Hungarian algorithm solves the \emph{linear assignment problem}, which in this setting may be expressed as
\begin{equation}
  \min_{\set{M}} \sum_{(i, j) \in \set{M}} d(\vec{x}_{T_{i}}, \vec{x}_{C_{j}}), \text{ such that } \forall (i, j), (k, l) \in \set{M}, i = k \iff j = l;
\end{equation}
the ``such that'' condition simply specifies that the matching is one-to-one. As such, optimal matching via the Hungarian algorithm minimizes the total sum of distances between matches in the matched subset.

\paragraph{Extensions.} The Hungarian algorithm can be extended to the $1:k$ matching case via $b$-matching while retaining optimality.
This process involves adding $k$ ``copy'' nodes for each node representing a $T$ datapoint, each with identical edges as the original node, and then perform the optimal matching on the resulting graph.
However, as~\textcite{khan_efficient_2016} notes, this can lead to calculations which quickly become time consuming as the total number of data points increases; the runtime is $O((k \cdot n + m)^3)$.~\textcite{khan_efficient_2016} goes on further to explore efficient approximation algorithms of the $b$-matching, such as the $b$-\textsc{Suitor} algorithm.
